251 MA

### 2022-12-01 code/process.Rmd
update NA counts

d |> filter(!status %in% c("missing", "non-experiment", "reproduction")) |> tally()
```
# Validation

@@

We parse out values from (for now) t tests, F tests, and chisq. 
Other stats to be added later, chisq ES to be added.

@@

parse_t <- function(tval, within_between) {
  df=str_extract(tval, "\\(.*\\)") |> str_sub(2,-2) |> as.numeric()
  val=str_extract(tval, "=.*") |> str_sub(2,-1) |> as.numeric() |> abs()
  pval=pt(q=val, df=df, lower.tail=FALSE)*2
  d_calc=NA
  if(within_between=="between"){
    d_calc=2*val/(sqrt(df))  # note this uses the 2t/sqrt(df) approximation df+2 might be more appropriate? 
  }
  else{d_calc=val/sqrt(df)}
  return(data.frame("df_1"=NA, "df_2"=df,"tstat"=val, "fstat"=NA, "p_calc"=pval, "d_calc"=d_calc))
}

parse_f <- function(fval, within_between){
  df_1 =str_extract(fval, "\\(.*,") |> str_sub(2,-2) |> as.numeric()
  df_2 =str_extract(fval, ",.*\\)") |> str_sub(2,-2) |> as.numeric()
  val=str_extract(fval, "=.*") |> str_sub(2,-1) |> as.numeric() |> abs()
  pval=pf(q=val, df1=df_1, df2=df_2, lower.tail=FALSE)
  d_calc=NA
  if(!is.na(df_1)&&df_1==1){
    if(within_between=="between"){
      d_calc=2*sqrt(val)/sqrt(df_2) # if there's just two groups, then F=t**2 and we can use the t-test thingy
    }
    else{
      d_calc=sqrt(val)/sqrt(df_2)
    }
  }
  else if(!is.na(df_1)){
    # we go through partial eta
    n=df_1+df_2+1 # rederive n b/c it's not always matching the n we have
    partial_eta=(val*df_1)/(val*df_1+df_2)
    d_calc=sqrt(n/(n-1)*(partial_eta/(1-partial_eta)))
  }
  return(data.frame("df_1"=df_1,"df_2"=df_2,"tstat"=NA, "fstat"=val, "p_calc"=pval, "d_calc"=d_calc))
}

parse_chisq <- function(chival, within_between){
  df=str_extract(chival, "\\(.*,") |> str_sub(2,-2) |> as.numeric()
  n=str_extract(chival, ",N=.*\\)") |> str_sub(4,-2) |> as.numeric()
  val=str_extract(chival, "\\)=.*") |> str_sub(3,-1) |> as.numeric() |> abs()
  print(val)
  pval=pchisq(val,df, lower.tail=F)
  phi=sqrt(val/n)
  d_calc=NA
  if (!is.na(df)&&df==1){
    d_calc=esc_phi(phi,pval,n, es.type="d")$es
  }
  return(data.frame("df_1"=df,"df_2"=NA,"tstat"=NA, "fstat"=NA, "p_calc"=pval, "d_calc"=d_calc))
}

#foo = "MCI m1=3.41[3.24,3.55], m2=3.44[3.24,3.57]"
parse_mean_ci <- function(mci,within_between,n){
  cond1=str_extract(mci, "m1.*\\],")
  cond2=str_extract(mci, "m2.*\\]")
  m1=str_extract(cond1, "m1=.*\\[") |> str_sub(4,-2) |> as.numeric()
  m2=str_extract(cond2, "m2=.*\\[") |> str_sub(4,-2) |> as.numeric()
  low1=str_extract(cond1, "\\[.*,.") |> str_sub(2,-3) |> as.numeric()
  low2=str_extract(cond2, "\\[.*,") |> str_sub(2,-2) |> as.numeric()
  high1=str_extract(cond1, ",.*\\]") |> str_sub(2,-2) |> as.numeric()
  high2=str_extract(cond2, ",.*\\]") |> str_sub(2,-2) |> as.numeric()
  se1=high1-low1/(2*1.96)
  se2=high2-low2/(2*1.96)
  se=(se1+se2)/2
  tval=(m1-m2)/se |> abs()
  if (within_between=="within"){
  df=n-1
  pval=pt(q=tval, df=df, lower.tail=FALSE)*2
  d_calc=tval/(sqrt(df)) 
  }
  else{
    df=NA
    pval=NA
    d_calc=NA
  }
  return(data.frame("df_1"=df,"df_2"=NA,"tstat"=tval, "fstat"=NA, "p_calc"=pval, "d_calc"=d_calc))
}

#parse_mean_ci(foo)

do_blanks <- function(){
  return(data.frame("df_1"=NA,"df_2"=NA,"tstat"=NA, "fstat"=NA, "p_calc"=NA, "d_calc"=NA))
}
do_parsing=function(raw_stat, within_between,n){
  if (is.na(raw_stat)) {return (do_blanks())}
  if (str_sub(raw_stat,1,1)=="t"){return(parse_t(raw_stat, within_between))}
  if (str_sub(raw_stat,1,1)=="F"){return(parse_f(raw_stat, within_between))}
  if (str_sub(raw_stat,1,1)=="c"){return(parse_chisq(raw_stat, within_between))}
  if (str_sub(raw_stat, 1,3)=="MCI"){return(parse_mean_ci(raw_stat, within_between,n))}
  return (do_blanks())
}
```

```{r}
fill_es <- function(type, value, n){
  if (is.na(type)) {
    return(NA)}
  if (is.na(value)) {
    return(NA)
  }
  if (type=="partial eta sq"){
    d = sqrt( (n-1)/n * abs(value)/(1-abs(value)))
    if (value < 0) {
      return(-d)}
    return(d)
    }
  if (type %in% c("d", "SMD")){
    return(abs(value))
  }
  return(NA)
}

@@

@@

```
Comparing calculated ES to original (or hand calculated) ES. All converted to SMD.

@@

## what to do with the weirdos

@@

```{r}
test |> filter(target_d_calc>1.9*d_fill_target)
test |> filter(target_d_calc<.6*d_fill_target)

test |> filter(rep_d_calc>1.9*d_fill_rep)
test |> filter(rep_d_calc<.6*d_fill_rep)

@@

@@

# Where does df not line up?
Note: it's not *necessarily* a problem if df's don't line up -- there could be some repeated measures or welch's test stuff going on, but nice to check.

test

@@

# Are the ES's okay?

@@

d_es |> ggplot(aes(x=target_d_calc, y=rep_d_calc, color=as.factor(replicated_report_code)))+
  geom_point()+
  coord_cartesian(xlim=c(0,2), ylim=c(-1,3))+
  geom_abline(slope=1, intercept=0)+
  geom_hline(yintercept=0)+
  scale_color_viridis(discrete=T, direction=-1)+
  labs(color="Replicated?", x="Original SMD", y="Replication SMD")
```

@@

```{r, message=T}
message("total rows")
d |> filter(!status %in% c("missing", "non-experiment", "reproduction")) |> nrow()

@@

d |> filter(!status %in% c("missing", "non-experiment", "reproduction")) |> summarize(                                                      across(everything(),~sum(is.na(.)))) |>
pivot_longer(everything()) |>
  arrange(desc(value)) |>
  filter(value!=0)

>>>>

d |> filter(!status %in% c("missing", "non-experiment", "reproduction")) |> tally()
```

@@

We parse out values from the raw stats.

@@

parse_t <- function(tval, within_between) {
  df=str_extract(tval, "\\(.*\\)") |> str_sub(2,-2) |> as.numeric()
  val=str_extract(tval, "=.*") |> str_sub(2,-1) |> as.numeric() |> abs()
  pval=pt(q=val, df=df, lower.tail=FALSE)*2
  d_calc=NA
  N_from_df=NA
  if(within_between=="between"){
    d_calc=2*val/(sqrt(df))  # note this uses the 2t/sqrt(df) approximation df+2 might be more appropriate? 
    N_from_df=df+2 # between so add 2
  }
  else{
    d_calc=val/sqrt(df)
    N_from_df=df+1
  }
  return(data.frame("df_1"=NA, "df_2"=df,"tstat"=val, "fstat"=NA, "p_calc"=pval, "d_calc"=d_calc, "N_calc"=N_from_df))
}
```

```{r}
parse_f <- function(fval, within_between){
  df_1 =str_extract(fval, "\\(.*,") |> str_sub(2,-2) |> as.numeric()
  df_2 =str_extract(fval, ",.*\\)") |> str_sub(2,-2) |> as.numeric()
  val=str_extract(fval, "=.*") |> str_sub(2,-1) |> as.numeric() |> abs()
  pval=pf(q=val, df1=df_1, df2=df_2, lower.tail=FALSE)
  d_calc=NA
  N_from_df=df_1+df_2+1
  if(!is.na(df_1)&&df_1==1){
    if(within_between=="between"){
      d_calc=2*sqrt(val)/sqrt(df_2) # if there's just two groups, then F=t**2 and we can use the t-test thingy
    }
    else{
      d_calc=sqrt(val)/sqrt(df_2)
    }
  }
  else if(!is.na(df_1)){
    # we go through partial eta
    n=df_1+df_2+1 # rederive n b/c it's not always matching the n we have
    partial_eta=(val*df_1)/(val*df_1+df_2)
    d_calc=sqrt(n/(n-1)*(partial_eta/(1-partial_eta)))
  }
  return(data.frame("df_1"=df_1,"df_2"=df_2,"tstat"=NA, "fstat"=val, "p_calc"=pval, "d_calc"=d_calc, "N_calc"=N_from_df))
}
```

```{r}
parse_chisq <- function(chival, within_between){
  df=str_extract(chival, "\\(.*,") |> str_sub(2,-2) |> as.numeric()
  n=str_extract(chival, ",N=.*\\)") |> str_sub(4,-2) |> as.numeric()
  val=str_extract(chival, "\\)=.*") |> str_sub(3,-1) |> as.numeric() |> abs()
  pval=pchisq(val,df, lower.tail=F)
  phi=sqrt(val/n)
  d_calc=NA
  if (!is.na(df)&&df==1){
    d_calc=esc_phi(phi,pval,n, es.type="d")$es
  }
  return(data.frame("df_1"=df,"df_2"=NA,"tstat"=NA, "fstat"=NA, "p_calc"=pval, "d_calc"=d_calc, "N_calc"=n))
}

```{r}
parse_mean_ci <- function(mci,within_between,n){
  cond1=str_extract(mci, "m1.*\\],")
  cond2=str_extract(mci, "m2.*\\]")
  m1=str_extract(cond1, "m1=.*\\[") |> str_sub(4,-2) |> as.numeric()
  m2=str_extract(cond2, "m2=.*\\[") |> str_sub(4,-2) |> as.numeric()
  low1=str_extract(cond1, "\\[.*,.") |> str_sub(2,-3) |> as.numeric()
  low2=str_extract(cond2, "\\[.*,") |> str_sub(2,-2) |> as.numeric()
  high1=str_extract(cond1, ",.*\\]") |> str_sub(2,-2) |> as.numeric()
  high2=str_extract(cond2, ",.*\\]") |> str_sub(2,-2) |> as.numeric()
  se1=high1-low1/(2*1.96)
  se2=high2-low2/(2*1.96)
  se=(se1+se2)/2
  tval=(m1-m2)/se |> abs()
  if (within_between=="within"){
  df=n-1
  pval=pt(q=tval, df=df, lower.tail=FALSE)*2
  d_calc=tval/(sqrt(df)) 
  }
  else{
    df=NA
    pval=NA
    d_calc=NA
  }
  return(data.frame("df_1"=df,"df_2"=NA,"tstat"=tval, "fstat"=NA, "p_calc"=pval, "d_calc"=d_calc, "N_calc"=NA))
}
```

```{r}
parse_mean_sd <- function(msd,within_between,n){
  cond1=str_extract(msd, "m1.*\\),")
  cond2=str_extract(msd, "m2.*\\)")
  m1=str_extract(cond1, "m1=.*\\(") |> str_sub(4,-2) |> as.numeric()
  m2=str_extract(cond2, "m2=.*\\(") |> str_sub(4,-2) |> as.numeric()
  sd1=str_extract(cond1, "\\(.*\\),") |> str_sub(2,-3) |> as.numeric()
  sd2=str_extract(cond2, "\\(.*\\)") |> str_sub(2,-2) |> as.numeric()
  sd_pool=(sd1+sd2)/2
  se=sd_pool/sqrt(n)
  tval=(m1-m2)/se |> abs()
  if (within_between=="within"){
  df=n-1
  pval=pt(q=tval, df=df, lower.tail=FALSE)*2
  d_calc=tval/(sqrt(df)) 
  }
  else if (within_between=="between"){
    df=n-2
    pval=pt(q=tval, df=df, lower.tail=FALSE)*2
    d_calc=2*tval/sqrt(df)
  }
  else{
    df=NA
    pval=NA
    d_calc=NA
  }
  return(data.frame("df_1"=df,"df_2"=NA,"tstat"=tval, "fstat"=NA, "p_calc"=pval, "d_calc"=d_calc, "N_calc"=NA))
}

```{r}
parse_mean_se <- function(mse,within_between,n){
  cond1=str_extract(mse, "m1.*\\),")
  cond2=str_extract(mse, "m2.*\\)")
  m1=str_extract(cond1, "m1=.*\\(") |> str_sub(4,-2) |> as.numeric()
  m2=str_extract(cond2, "m2=.*\\(") |> str_sub(4,-2) |> as.numeric()
  se1=str_extract(cond1, "\\(.*\\),") |> str_sub(2,-3) |> as.numeric()
  se2=str_extract(cond2, "\\(.*\\)") |> str_sub(2,-2) |> as.numeric()
  se=(se1+se2)/2
  tval=(m1-m2)/se |> abs()
  if (within_between=="within"){
  df=n-1
  pval=pt(q=tval, df=df, lower.tail=FALSE)*2
  d_calc=tval/(sqrt(df)) 
  }
  else if (within_between=="between"){
    df=n-2
    pval=pt(q=tval, df=df, lower.tail=FALSE)*2
    d_calc=2*tval/sqrt(df)
  }
  else{
    df=NA
    pval=NA
    d_calc=NA
  }
  return(data.frame("df_1"=df,"df_2"=NA,"tstat"=tval, "fstat"=NA, "p_calc"=pval, "d_calc"=d_calc, "N_calc"=NA))
}

```{r}
parse_wilcoxon <- function(wilcox){
  z=str_sub(wilcox,11,-1) |> as.numeric()
  pval=pnorm(z,lower.tail=F)
  #r=z/sqrt(n)
  #d_calc=4*r/(sqrt(1-r**2))
  d_calc=NA
    return(data.frame("df_1"=NA,"df_2"=NA,"tstat"=NA, "fstat"=NA, "p_calc"=pval, "d_calc"=d_calc, "N_calc"=NA))

}
```

do_blanks <- function(){
  return(data.frame("df_1"=NA,"df_2"=NA,"tstat"=NA, "fstat"=NA, "p_calc"=NA, "d_calc"=NA, "N_calc"=NA))
}
do_parsing=function(raw_stat, within_between,n){
  if (is.na(raw_stat)) {return (do_blanks())}
  if (str_sub(raw_stat,1,1)=="t"){return(parse_t(raw_stat, within_between))}
  if (str_sub(raw_stat,1,1)=="F"){return(parse_f(raw_stat, within_between))}
  if (str_sub(raw_stat,1,1)=="c"){return(parse_chisq(raw_stat, within_between))}
  if (str_sub(raw_stat, 1,3)=="MCI"){return(parse_mean_ci(raw_stat, within_between,n))}
  if (str_sub(raw_stat, 1,3)=="MSD"){return(parse_mean_sd(raw_stat, within_between,n))}
  if (str_sub(raw_stat, 1,3)=="MSE"){return(parse_mean_se(raw_stat, within_between,n))}
  if (str_sub(raw_stat, 1,8)=="wilcoxon"){return(parse_wilcoxon(raw_stat))}

  return (do_blanks())
}
```

```{r}
fill_es <- function(type, value, n){
  if (is.na(type)) {
    return(NA)}
  if (is.na(value)) {
    return(NA)
  }
  if (type=="partial eta sq"){
    d = sqrt( (n-1)/n * abs(value)/(1-abs(value)))
    return(d)
    }
  if (type %in% c("d", "SMD")){
    return(abs(value))
  }
  return(NA)
}

@@

## Compare pvals

@@

## Compare ES

Comparing calculated ES to original (or hand calculated) ES. All converted to SMD.

@@

@@

@@

these are ones where we did not get a non-NA effect size. Could be due to formulas being incomplete, or missing information.

@@

## how are the df's lining up with sample size?

We can try to back out effective sample size from the df's. Note that we don't always expect them to agree (Welch's t test, repeated measures, comparisons only on a sample), but still useful to look at.

There's a lot of off by 1 or 2 things, which I'm unsure what to make of. (As much as possible both sample size and test stats are taken from the papers directly)

```{r}
 test |> filter(!is.na(target_N_calc)) |> select(target_N_calc, target_N, target_raw_stat, within_between, target_lastauthor_year) |>
  mutate(diff=target_N_calc-target_N) |> 
  ggplot(aes(x=target_N, y=target_N_calc, color=(target_N==target_N_calc)))+geom_point()+geom_abline(slope=1, intercept=0)+scale_y_log10()+scale_x_log10()+scale_color_manual(values=c("TRUE"="black","FALSE"="red"))

 test |> filter(!is.na(rep_N_calc)) |> select(rep_N_calc, replication_N, replication_raw_stat, within_between, target_lastauthor_year) |> mutate(diff=rep_N_calc-replication_N) |> 
   ggplot(aes(x=replication_N, y=rep_N_calc, color=(replication_N==rep_N_calc)))+geom_point()+geom_abline(slope=1, intercept=0)+scale_y_log10()+scale_x_log10()+scale_color_manual(values=c("TRUE"="black","FALSE"="red"))

@@

@@

d_es |> ggplot(aes(x=target_d_calc, y=rep_d_calc, color=as.factor(replicated_report_code)))+
  geom_point()+
  coord_cartesian(xlim=c(0,5), ylim=c(-1,4))+
  geom_abline(slope=1, intercept=0)+
  geom_hline(yintercept=0)+
  scale_color_viridis(discrete=T, direction=-1)+
  labs(color="Replicated?", x="Original SMD", y="Replication SMD")
```

@@

Note, it may seem weird that we're missing d and SE for many more replications than we are p values. This is because we can't get d_calc if we don't have a filled in value for same direction (but we have p value and unsigned d_calc).

```{r, message=T}
message("total rows")
d_es |> filter(!status %in% c("missing", "non-experiment", "reproduction")) |> nrow()

@@

d_es |> filter(!status %in% c("missing", "non-experiment", "reproduction")) |> ungroup() |> summarize(                                                      across(everything(),~sum(is.na(.)))) |>
  select(-rep_fstat, -rep_tstat,-rep_df_1,-rep_df_2,
         -target_fstat, -target_tstat, -target_df_1, -target_df_2,
         -d_fill_rep, -d_fill_target, 
         -replication_ES_type, -replication_ES_value, -replication_pvalue,
         -target_ES_type, -target_ES_value, -target_pvalue,
         -rep_N_calc, -target_N_calc,
         -replication_test, -target_test) |> 
pivot_longer(everything()) |>
  arrange(desc(value)) |>
  filter(value!=0)
